---
title: "Data Import"
output: github_document
---

This file is for doing data import

```{r}
library(tidyverse)
library(readxl)
library(haven)
```

## Importing two datasets

Import out first dataset
```{r}
litters_df =
  read_csv(
    file = "data_file/FAS_litters.csv",
   na = c("", "NA",".")
  )
litters_df

# makes it into snake case 
litters_df =janitor:: clean_names(litters_df)
```
Import the pups dataset
using na formats all the missing values (.) and the NAs to read as missing 
```{r}
pups_df = 
  read_csv(
    file = "data_file/FAS_pups.csv", 
    skip = 3,
    na = c("","NA",".")
  )
  
pups_df = janitor::clean_names(pups_df) 
  
```

## Look at the data 
```{r}
litters_df
pups_df
```

Head or tail of the dataset 
```{r}
head(litters_df, 20)

tail(pups_df, 5)
```

skimming? 
using :: is for when you want to use one feature of a package without loading the whole package. 
```{r}
skimr::skim(pups_df)
```

viewing- don't do this in R markdown just in the command 
```{r}
# view(pups_df) 
```

tells you what you can do when reading a csv
```{r}
#?read_csv
```

## oh excel .....
Importing new dataset from excel 
```{r}
mlb_df =
  read_excel("data_file/mlb11.xlsx")
```

Look at the data 
```{r}
mlb_df

```

tells you what you can do when reading an excel file
```{r}
# ?read_excel 
```

Import FOTR words
```{r}
FOTR_df= 
  read_excel("data_file/LotR_Words.xlsx",
              range = "B3:D6"
             )

FOTR_df

```

Importing the Two Towers Table 
```{r}
FOTR2_df= 
  read_excel("data_file/LotR_Words.xlsx",
              range = "F3:H6"
             )
FOTR2_df
```

## Import SAS 
Read in the pulse dataset
```{r}
Pulse_df=
  read_sas("data_file/public_pulse_data.sas7bdat")

Pulse_df = janitor::clean_names(Pulse_df)

```

## What's so bad about 'read.csv()'
Import pups with 'read.csv()
```{r}
pups_df_base =
  read.csv("data_file/FAS_pups.csv")

pups_df_base
```

Won't show: lazy about variable names; super slow for big datasets 

