---
title: "Data Import"
output: github_document
---

This file is for doing data import

```{r}
library(tidyverse)
library(readxl)
library(haven)
```

## Importing two datasets

Import out first dataset
```{r}
litters_df =
  read_csv(
    file = "data_file/FAS_litters.csv",
   na = c("", "NA",".")
  )
litters_df

# makes it into snake case 
litters_df =janitor:: clean_names(litters_df)
```

Import the pups dataset
using na formats all the missing values (.) and the NAs to read as missing 
```{r}
pups_df = 
  read_csv(
    file = "data_file/FAS_pups.csv", 
    skip = 3,
    na = c("","NA",".")
  )
  
pups_df = janitor::clean_names(pups_df) 
  
```

### 'select' some variables
```{r}
# shows only the variable you name 
select(litters_df, group, litter_number)

select(litters_df, group:gd_of_birth)

select(litters_df, -group)

select(litters_df, starts_with("gd"), group)

# this renames the variable to GROUP and then keeps everything else
select(litters_df, GROUP = group, everything())

select(litters_df, -starts_with("gd"))

```

a couple of alternatives 

```{r}
# another way to rename variables 
rename(litters_df, GROUP = group)

# moves these variables to the front
relocate(litters_df, litter_number, gd0_weight)
```

In the pups data, select litter number, sex, and PD ears
```{r}
# shows only the variable you name 
select(pups_df, litter_number, sex, pd_ears)
```

## 'filter' some rows 
```{r}
filter(litters_df, gd_of_birth == 20) 

filter(litters_df, pups_born_alive <= 5)

filter(litters_df, pups_born_alive != 6)

filter(litters_df, group %in% c("Con7", "Con8"), pups_born_alive < 7) 

# removes the rows with the missing values for those variables 
drop_na(litters_df, gd0_weight) 
```

In the pups dataset, keep only rows where PD walk is less than 11 and sex is 2
```{r}
filter(pups_df, pd_walk < 11, sex == 2)
```

## 'mutate' to change some columns 

lets add and change some columns 
str_to_lower = makes everything lower case 
```{r}
mutate(litters_df, 
       wt_gain= gd18_weight - gd0_weight,
       wt_gain_squared = wt_gain ^ 2, 
       group = str_to_lower(group))
```

In the pups dataset, create a variable that subtracts 7 from PD walk

```{r}
mutate(pups_df, pd_walk_minus_7 = pd_walk - 7)
```

## 'arange' the data 
```{r}
arrange(litters_df, pups_born_alive)
arrange(litters_df, desc(pups_born_alive))

```

Good example
pipe is shift command m 
```{r}
litters_df =
  read_csv(file = "data_file/FAS_litters.csv",na = c("", "NA",".")) |> 
  janitor:: clean_names() |> 
  select(group, starts_with("gd")) |> 
    drop_na() |> 
    mutate(
      wt_gain = gd18_weight - gd0_weight,
      group = str_to_lower(group)
    )
```

load pups, clean names, drop missing, keep litter number and pd variables, add pd walk -7
```{r}
pups_df =
  read_csv(file = "data_file/FAS_pups.csv", skip = 3, na = c("", "NA",".")) |> 
  janitor:: clean_names() |> 
  drop_na() |> 
  select(litter_number, starts_with("pd")) |> 
    mutate(
      pd_walk_minus_7 = pd_walk - 7
    )
```

sometimes the first thing isn't a dataframe 

```{r}
litters_df |> 
  filter(group %in% c("con7", "con8" )) |> 
  lm(gd18_weight ~ gd0_weight, data = _)
```

select vs pull 
```{r}
litters_df |> 
  select(group)

litters_df |> 
  pull(group) 
```

never do '$'
```{r}
litters_df$group
```

pipes and viewing (do not usually have this in an R markdown)
```{r}
litters_df |> 
  filter(group %in% c("con7", "con8"))
```

